---
title: "Demo"
output: html_document
date: "2024-03-16"
editor_options: 
  chunk_output_type: inline
---

```{r }
library(dplyr)
library(lubridate)
library(ggplot2)
library(readr)
library(RSQLite)
library(DBI)
library(readxl)
library(tidyr)
library(plotly)
```

```{r}
db <- RSQLite::dbConnect(RSQLite::SQLite(), "e_commerce_database")
```


```{sql connection=db}
CREATE TABLE IF NOT EXISTS Supplier (
  SUPPLIER_ID VARCHAR(255) PRIMARY KEY NOT NULL,
  SUPPLIER_NAME VARCHAR(255) NOT NULL,
  SUPPLIER_PHONE VARCHAR(255),
  SUPPLIER_EMAIL VARCHAR(255)
);

CREATE TABLE IF NOT EXISTS Product (
  PRODUCT_ID VARCHAR(255) PRIMARY KEY NOT NULL,
  PRODUCT_NAME VARCHAR(255) NOT NULL,
  PRODUCT_CATEGORY VARCHAR(255),
  PRICE FLOAT NOT NULL,
  SUPPLIER_ID VARCHAR(255) NOT NULL,
  FOREIGN KEY(SUPPLIER_ID) REFERENCES Supplier(SUPPLIER_ID)
);

CREATE TABLE IF NOT EXISTS Inventory (
  INVENTORY_ID VARCHAR(255) PRIMARY KEY NOT NULL,
  STOCK INTEGER NOT NULL,
  SHELF_NO VARCHAR(255),
  PRODUCT_ID VARCHAR(255) NOT NULL,
  FOREIGN KEY(PRODUCT_ID) REFERENCES Product(PRODUCT_ID)
);

CREATE TABLE IF NOT EXISTS Customer (
  CUSTOMER_ID VARCHAR(255) PRIMARY KEY NOT NULL,
  CUSTOMER_FIRSTNAME VARCHAR(255) NOT NULL,
  CUSTOMER_LASTNAME VARCHAR(255) NOT NULL,
  CUSTOMER_EMAIL VARCHAR(255),
  CUSTOMER_PHONE VARCHAR(255),
  CUSTOMER_BIRTHDAY DATE,
  CUSTOMER_GENDER VARCHAR(50),
  SHIPMENT_ID VARCHAR(255),
  PAYMENT_ID VARCHAR(255)
);

CREATE TABLE IF NOT EXISTS Shipment (
  SHIPMENT_ID VARCHAR(255) PRIMARY KEY NOT NULL,
  SHIPMENT_DATE DATE NOT NULL,
  SHIPMENT_ADDRESS VARCHAR(255) NOT NULL,
  SHIPMENT_CITY VARCHAR(255) NOT NULL,
  SHIPMENT_ZIPCODE VARCHAR(255) NOT NULL,
  BILLING_COUNTRY VARCHAR(255) NOT NULL,
  CUSTOMER_ID VARCHAR(255) NOT NULL,
  PRODUCT_ID VARCHAR(255) NOT NULL,
  FOREIGN KEY(CUSTOMER_ID) REFERENCES Customer(CUSTOMER_ID),
  FOREIGN KEY(PRODUCT_ID) REFERENCES Product(PRODUCT_ID)
);

CREATE TABLE IF NOT EXISTS Payment (
  PAYMENT_ID VARCHAR(255) PRIMARY KEY NOT NULL,
  PAYMENT_METHOD VARCHAR(255) NOT NULL,
  ORDER_AMOUNT FLOAT NOT NULL,
  PAYMENT_DATE DATE NOT NULL,
  BILLING_ADDRESS VARCHAR(255) NOT NULL,
  BILLING_CITY VARCHAR(255) NOT NULL,
  BILLING_ZIPCODE VARCHAR(255) NOT NULL,
  BILLING_COUNTRY VARCHAR(255) NOT NULL,
  CUSTOMER_ID VARCHAR(255) NOT NULL,
  PRODUCT_ID VARCHAR(255) NOT NULL,
  FOREIGN KEY(CUSTOMER_ID) REFERENCES Customer(CUSTOMER_ID),
  FOREIGN KEY(PRODUCT_ID) REFERENCES Product(PRODUCT_ID)
);

```

```{r}
suppliers_data <- read_csv('supplier_ecommerce.csv')
products_data <- read_csv('products_ecommerce.csv')
inventories_data <- read_csv('inventory_ecommerce.csv')
customers_data <- read_csv('customers_ecommerce.csv')
shipments_data <- read_csv('shipment_ecommerce.csv')
payments_data <- read_csv('payments_ecommerce.csv')
```

```{r}
dbWriteTable(db, "Supplier", suppliers_data, overwrite = TRUE)
dbWriteTable(db, "Product", products_data, overwrite = TRUE)
dbWriteTable(db, "Inventory", inventories_data, overwrite = TRUE)
dbWriteTable(db, "Customer", customers_data, overwrite = TRUE)
dbWriteTable(db, "Shipment", shipments_data, overwrite = TRUE)
dbWriteTable(db, "Payment", payments_data, overwrite = TRUE)
```

```{r}
# List all tables
tables <- dbListTables(db)
print(tables)
```
# lets start doing querying from easy to a bit complicated ones
```{sql connection=db}
SELECT * FROM CUSTOMER;
```

# lets try specific columns
```{sql connection=db}
SELECT CUSTOMER_ID, CUSTOMER_FIRSTNAME, CUSTOMER_LASTNAME FROM CUSTOMER;
```
# Lets see PRODUCTS and their prices
```{sql connection=db}
SELECT PRODUCT_NAME, PRICE FROM PRODUCT;
```
# lets use wildcard expression and Limit with it
```{sql connection=db}
SELECT * FROM SUPPLIER LIMIT 5;
```
# our top five expensive items
```{sql connection=db}
SELECT DISTINCT PRODUCT_ID, PRODUCT_NAME, PRICE
FROM Product
WHERE PRICE > 1000
ORDER BY PRICE DESC
LIMIT 5;
```
# our five cheapeast items
```{sql connection=db}
SELECT DISTINCT PRODUCT_ID, PRODUCT_NAME, PRICE
FROM Product
ORDER BY PRICE ASC
LIMIT 5;
```

TOTAL SALES
```{sql connection=db}
SELECT SUM(ORDER_AMOUNT) AS Total_Sales FROM Payment;

```
# Most sold item with quantity sold, price per item, total quantity sold, and total profit:
```{sql connection=db}

SELECT p.PRODUCT_NAME, p.PRICE, COUNT(p.PRODUCT_ID) AS Units_Sold, SUM(pay.ORDER_AMOUNT) AS Total_Sales
FROM Product p
JOIN Payment pay ON p.PRODUCT_ID = pay.PRODUCT_ID
GROUP BY p.PRODUCT_ID
ORDER BY Units_Sold DESC, Total_Sales DESC
LIMIT 1;

```
# For the category with the most shipments
```{sql connection=db}
SELECT p.PRODUCT_CATEGORY, COUNT(s.SHIPMENT_ID) AS Total_Shipments
FROM Product p
JOIN ShipMENT s ON p.PRODUCT_ID = s.PRODUCT_ID
GROUP BY p.PRODUCT_CATEGORY
ORDER BY Total_Shipments DESC
LIMIT 1;

```
# For the category with the least shipments
```{sql connection=db}
SELECT p.PRODUCT_CATEGORY, COUNT(s.SHIPMENT_ID) AS Total_Shipments
FROM Product p
JOIN ShipMENT s ON p.PRODUCT_ID = s.PRODUCT_ID
GROUP BY p.PRODUCT_CATEGORY
ORDER BY Total_Shipments ASC
LIMIT 1;


```
# Customers with the Highest Spending
```{sql connection=db}
SELECT c.CUSTOMER_FIRSTNAME, c.CUSTOMER_LASTNAME, p.PRODUCT_NAME, SUM(pa.ORDER_AMOUNT) AS Total_Spent
FROM Customer c
JOIN Payment pa ON c.CUSTOMER_ID = pa.CUSTOMER_ID
JOIN Product p ON pa.PRODUCT_ID = p.PRODUCT_ID
GROUP BY c.CUSTOMER_ID, p.PRODUCT_NAME
ORDER BY Total_Spent DESC
LIMIT 10;

```

# Inventory Levels Below Reorder Threshold Including Product Names
```{sql connection=db}

SELECT p.PRODUCT_NAME, i.STOCK
FROM Inventory i
JOIN Product p ON i.PRODUCT_ID = p.PRODUCT_ID
WHERE i.STOCK < 10; -- Assuming 10 is the reorder threshold


```

# Advanced data analysis


# Monthly sales over time
```{r}
str(payments_data)

payments_data$PAYMENT_DATE <- as.Date(payments_data$PAYMENT_DATE, format = "%m/%d/%Y") # Adjust format as needed

sales_data <- payments_data %>%
  mutate(MONTH = floor_date(PAYMENT_DATE, "month")) %>%
  group_by(MONTH) %>%
  summarise(Total_Sales = sum(ORDER_AMOUNT))



plot_ly(data = sales_data, x = ~MONTH, y = ~Total_Sales, type = 'scatter', mode = 'lines+markers',
        line = list(shape = "spline", smoothing = 1.3), marker = list(size = 10)) %>%
  layout(title = "Monthly Sales Over Time",
         xaxis = list(title = "Month",
                      tickformat = "%b %Y",
                      type = "date"),
         yaxis = list(title = "Total Sales ($)"),
         hovermode = "closest")

```




# volume sold over the months
```{r}
sales_volume <- payments_data %>%
  mutate(MONTH = floor_date(as.Date(PAYMENT_DATE), "month")) %>%
  group_by(MONTH) %>%
  summarise(Sales_Volume = n())

plot_ly(data = sales_volume, x = ~MONTH, y = ~Sales_Volume, type = 'scatter', mode = 'lines+markers',
        line = list(shape = "spline"), marker = list(size = 10)) %>%
  layout(title = "Monthly Sales Volume",
         xaxis = list(title = "Month",
                      tickformat = "%b %Y",
                      type = "date"),
         yaxis = list(title = "Number of Items Sold"))

```




```{r}

sales_category_data <- payments_data %>%
  left_join(products_data, by = "PRODUCT_ID") %>%
  mutate(MONTH = floor_date(as.Date(PAYMENT_DATE), "month")) %>%
  group_by(MONTH, PRODUCT_CATEGORY) %>%
  summarise(Total_Sales = sum(ORDER_AMOUNT), .groups = 'drop')

plot_ly(data = sales_category_data) %>%
  add_bars(x = ~MONTH, y = ~Total_Sales, color = ~PRODUCT_CATEGORY, colors = "Pastel1") %>%
  layout(title = "Monthly Sales Revenue by Product Category",
         barmode = 'group',
         xaxis = list(title = "Month", tickformat = "%b %Y", type = "date"),
         yaxis = list(title = "Total Sales ($)"),
         colorway = RColorBrewer::brewer.pal(8, "Dark2"))



```


```{r}

str(customers_data)

# Convert dates to Date objects, assuming the format is month-day-year
customers_data$CUSTOMER_BIRTHDAY <- mdy(customers_data$CUSTOMER_BIRTHDAY)

# Now we can safely extract the year
customers_data$BirthYear <- year(customers_data$CUSTOMER_BIRTHDAY)



customers_data <- customers_data %>%
  mutate(Generation = case_when(
    BirthYear >= 1928 & BirthYear <= 1945 ~ "Silent Generation",
    BirthYear >= 1946 & BirthYear <= 1964 ~ "Baby Boomers",
    BirthYear >= 1965 & BirthYear <= 1980 ~ "Generation X",
    BirthYear >= 1981 & BirthYear <= 1996 ~ "Millennials",
    BirthYear >= 1997 & BirthYear <= 2012 ~ "Generation Z",
    BirthYear >= 2013 ~ "Generation Alpha",
    TRUE ~ "Unknown" # for any data that doesn't fit the above categories
  ))


ggplot(customers_data, aes(x = Generation)) +
  geom_bar(fill = "cornflowerblue") +
  labs(title = "Distribution of Customers by Generation",
       x = "Generation",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

library(plotly)

plot_ly(customers_data, x = ~Generation, type = 'histogram', marker = list(color = 'lightsalmon')) %>%
  layout(title = "Distribution of Customers by Generation",
         xaxis = list(title = "Generation"),
         yaxis = list(title = "Count"))

```
```{r}


```

